Data Exploration: IUCN and AquaMaps

### 2015-11-26 data_explore.Rmd
notes added to manuscript; tried a map of difference in frequency between the two datasets but that still needs work

```{r, echo = FALSE, warning = FALSE, message = FALSE}

@@

```{r get spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
data_file <- 'data/spp_list_w_area_percent_noextinct.csv'

@@

``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
```

@@

```{r pct small within large, echo = FALSE, warning = FALSE, eval = TRUE}

@@

``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

@@

```{r map, echo=F, warning=F, message=F}

# Number of aquamaps species per cell

am_n_per_cell = filter(am_spp_cells,am_sid %in% unique(spp_list$am_sid))%>%
                  group_by(loiczid)%>%
                   summarize(count.am=n())%>%
                    as.data.frame()

am_sp_map = subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count.am')], 
                     by = 'loiczid', 
                     which = 'count.am', 
                     subsWithNA = TRUE)

#number of IUCN species per cell
iucn_n_per_cell = iucn_spp_cells%>%
                   group_by(loiczid)%>%
                    summarize(count.iucn=n())%>%
                    as.data.frame()

iucn_sp_map = subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count.iucn')], 
                     by = 'loiczid', 
                     which = 'count.iucn', 
                     subsWithNA = TRUE)

#Plot both maps
myTheme = rasterTheme(region = brewer.pal('Blues', n = 9))

@@

levelplot(iucn_sp_map, par.settings = myTheme, main = 'IUCN Dataset', colorkey = list(at = seq(0, 1063, length = 15)))
```

@@

```{r all aquamaps,eval=F, echo=F}

am_spp_cells_big = fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv'))
#am_spp_cell_big is missing loiczid info so bring in hcaf to match

@@

all_am = am_spp_cells_big %>%
          select(SpeciesID,CsquareCode) %>%
          group_by(CsquareCode) %>%
          summarize(count=n()) %>%
          as.data.frame()

all = left_join(all_am, hcaf)

all_am_map = subs(loiczid_raster, all[ , c('loiczid', 'count')], 
                     by = 'loiczid', 
                     which = 'count', 
                     subsWithNA = TRUE)

@@

![image](./figures/all_aquamaps_map.png)

>>>>

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

@@

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
data_file <- 'data/spp_list_w_area_percent_noextinct.csv'

@@

``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
# bar chart: number of spp per group
# scatter plot: area compare IUCN vs AM, including log plot and plot by species group
# rangemap: Chromis alpha 
```

@@

```{r pct_small_within_large, echo = FALSE, warning = FALSE, eval = TRUE}

@@

``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

@@

```{r global_dist_map, echo = FALSE, warning = FALSE, message = FALSE}

### find total species in each dataset for later percent calcs
am_n_tot   <- am_spp_cells %>% 
  filter(am_sid %in% unique(spp_list$am_sid)) %>%
  select(am_sid) %>% 
  unique() %>% nrow()
iucn_n_tot <- iucn_spp_cells %>% 
  filter(sciname %in% unique(spp_list$sciname)) %>%
  select(sciname) %>% 
  unique() %>% nrow()
### good - the same number

### Number of aquamaps species per cell
am_n_per_cell <- am_spp_cells %>%
  filter(am_sid %in% unique(spp_list$am_sid)) %>%
  group_by(loiczid) %>%
  summarize(count_am = n(),
            pct_am = count_am/am_n_tot) %>%
  as.data.frame()

am_sp_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count_am')], 
                  by = 'loiczid', 
                  which = 'count_am', 
                  subsWithNA = TRUE)

### Number of IUCN species per cell
iucn_n_per_cell <- iucn_spp_cells %>%
  group_by(loiczid) %>%
  summarize(count_iucn = n(),
            pct_iucn = count_iucn/iucn_n_tot) %>%
  as.data.frame()

iucn_sp_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count_iucn')], 
                    by = 'loiczid', 
                    which = 'count_iucn', 
                    subsWithNA = TRUE)

### Plot both maps
myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))

@@

levelplot(iucn_sp_map, par.settings = myTheme, 
          main = 'IUCN Dataset', 
          colorkey = list(at = seq(0, 1063, length = 15)))

@@

```{r all_aquamaps, eval = FALSE, echo = FALSE}

am_spp_cells_big <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv')) %>%
  as.data.frame()
### am_spp_cell_big is missing loiczid info so bring in hcaf to match

@@

### find total spp to allow for percent calcs later
am_all_n_tot <- am_spp_cells_big %>% 
  select(SpeciesID) %>% 
  unique() %>% nrow()

all_am <- am_spp_cells_big %>%
  select(SpeciesID, CsquareCode) %>%
  group_by(CsquareCode) %>%
  summarize(all_count = n(),
            all_pct   = all_count/am_all_n_tot) %>%
  as.data.frame()

all <- all_am %>%
  left_join(hcaf, by = 'CsquareCode')

all_am_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_count')], 
                   by = 'loiczid', 
                   which = 'all_count', 
                   subsWithNA = TRUE)

@@

![image](figures/all_aquamaps_map.png)

``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
### find distribution by percent of total species
am_sp_pct_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'pct_am')], 
                      by = 'loiczid', 
                      which = 'pct_am', 
                      subsWithNA = TRUE)
iucn_sp_pct_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'pct_iucn')], 
                        by = 'loiczid', 
                        which = 'pct_iucn', 
                        subsWithNA = TRUE)
### find difference in percent between the two distribution maps; should be small,
### if species are well-matched, since it's all the same species
diff_pct_map    <- iucn_sp_pct_map - am_sp_pct_map

### plot maps with distributions as percentages.  Should it be 0 - 1, or 
### 0 - max value?
levelplot(am_sp_pct_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset pct', 
          colorkey = list( at = seq(0, 1, length = 15)))
levelplot(iucn_sp_pct_map, par.settings = myTheme, 
          main = 'IUCN Dataset pct', 
          colorkey = list(at = seq(0, 1, length = 15)))
### this is the difference one. Should it be min value to max value?
levelplot(diff_pct_map, par.settings = myTheme, 
          main = 'IUCN pct - AM pct', 
          colorkey = list(at = seq(-1, 1, length = 15)))

### Do similar, but with the full aquamaps list (should also do this
### against the full IUCN list...)
all_am_pct_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_pct')], 
                   by = 'loiczid', 
                   which = 'all_pct', 
                   subsWithNA = TRUE)

diff_pct_map2    <- iucn_sp_pct_map - all_am_pct_map
levelplot(all_am_pct_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset, pct')
levelplot(diff_pct_map2, par.settings = myTheme, main = 'iucn (overlap only) - aquamaps (all)')
